A Framework for Discrete Change

### 2008-11-03 papers/individual/individual.tex
-added prelim results for WPT
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/depmix@235 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

Alternative non-strategy based analyses of the WPT (Lagnado et al, Speekenbrink et al) 
have estimated response strategies by variations of logistic regression.

%associative, rule-based.  
Here, we analyze the behavior of a single individual performing the
WPT for 200 trials. We let each state be characterized by a Generalized
Linear Model with a Bernoulli response and logistic link function. We are
interested in whether a DMM can recover a strategy model in line with Gluck et al. 
As we fit the data to a single subject, we must place some constraints.
Specifically, we constrain the state transitions to be in a ``left-right'' format
(states can only proceed to the immediately adjacent state and never back) and 
the initial state

A single state model (usual GLM) 
We started with a 3-state model,

>>>>

Alternative non-strategy based analyses of the WPT (Lagnado et al, Speekenbrink et al) 
have estimated response strategies by logistic regression, allowing the regression
coefficients to change over time.

%associative, rule-based.  
Here, we analyze the behavior of a single individual performing the
WPT for 200 trials. We chose to analyse the ``average'' participant (the
participant with performance closest to the group average) in a large
unpublished dataset. We let each state be characterized by a GLM with a 
Binomial distributed response and logistic link function (i.e., a logistic
regression model). We are particularly interested in evidence for 
strategy switching and whether a DMM can recover a strategy model 
in line with Gluck et al. (2002).

As we fit the data to a single subject, 
we must place some constraints. Specifically, we constrain the state 
transitions to be in a ``left-right'' format (states can only proceed 
to the immediately adjacent state and never back, and must start in 
the initial state). We fitted a single, two and three state model to 
the data. This showed that a two state model was better than a single
and three state model.

\begin{table}
\caption{Estimates for the weather prediction task}
\label{tab:WPT}
\begin{tabular}{lcccccccc} \hline
 & & \multicolumn{1}{c}{1 state} & & \multicolumn{2}{c}{2 state} && \multicolumn{2}{c}{2 state (constr.)} \\ \cline{3-3} \cline{5-6} \cline{8-9}
parameter & & $S_1$ & & $S_1$ & $S_2$ & & $S_1$ & $S_2$ \\ \hline
(intercept) & & -0.69 & & -2.73 & 0.88 & & -1.24 & 0 \\
cue 1 && 1.69 && 2.12 & 1.60 && 1.65 & 1.97 \\
cue 2 && 1.12 && 0.97 & 1.63 && 0 & 1.92 \\
cue 3 && -0.49 && 0.91 & -2.03 && 0 & -1.58 \\
cue 4 && -1.32 && 0.69 & -3.16 && 0 & -2.67 \\ \hline
 & & \multicolumn{1}{c}{AIC=204.47} & & \multicolumn{2}{c}{AIC=187.50} && \multicolumn{2}{c}{AIC=185.24}
\end{tabular}
\end{table}

Investigation of the parameter estimates (see Table~\ref{tab:WPT}) indicated that the first state might be
a single cue strategy (the regression coefficient for the first cue was of much 
larger magnitude than that of the other cues). The second state was a multi-cue
strategy (all cues had regression coefficients of reasonable magnitude). 
To reduce the degrees of freedom, and improve parameter estimates, we implemented 
constraints to force state 1 into a single cue strategy (fixing the coefficients
of the remaining three cues to 0) and state 2 in a multi-cue strategy (forcing
the intercept to 0). These restrictions resulted in a better AIC value of AIC=185.24 
(df=7). Interestingly, the single cue strategy was somewhat different than 
described by Gluck et al. Parameter estimates indicated relatively more consistent 
predictions of ``rain'' in the absence of cue 1 ($Pr(\text{sun}) = 0.22$) and more
inconsistent predictions of ``sun'' in the presence of cue 1 ($Pr(\text{sun}) = 0.60$). 
The cue weights of the multi-cue strategy were in the direction of the optimal weights. 
The Viterbi state sequence indicated that the participant used the single
cue strategy for the first 60 trials, and then switched to the multi-cue strategy.
